\documentclass[11pt, a4paper]{article}

% Helvetica
\usepackage[scaled]{helvet}
\renewcommand\familydefault{\sfdefault} 
\usepackage[T1]{fontenc}

% 2cm margins
\usepackage[margin=2cm]{geometry}

\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[center]{caption}
\usepackage{hyperref}
\usepackage{placeins}
\usepackage{booktabs}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\usepackage{threeparttable}
%\newcolumntype{L}{>{\centering\arraybackslash}m{3cm}}


\usepackage{apacite} 
%\usepackage{setspace}
%\doublespacing

\definecolor{green}{rgb}{0,1,0}
\newcommand{\NS}[1] {{\textcolor{green}{#1}}}
\newcommand{\TM}[1] {{\textcolor{orange}{#1}}}
\newcommand{\AT}[1] {{\textcolor{blue}{#1}}}


\begin{document}
\title{ESRC SDAI Application - Understanding domestic abuse using Big Data}
\date{}
\maketitle

\section{Introduction}

Domestic abuse is a complex phenomenon affecting people from all walks of life. It is increasingly recognised as a major public policy concern in many countries, including the UK \cite{ep}. While anyone can become a victim of domestic abuse, women are disproportionately affected, with more than 25\% of women, and 15\% of men in England and Wales reported to have experienced some form of domestic abuse since the age of 16 \cite{ONS}.  The current legal definition in the UK aims to capture the multifaceted nature of domestic abuse, by recognising that domestic abuse encompasses a wide range of behaviours, including emotional, sexual, and physical abuse, threatening, intimidating, coercive and controlling behaviour \cite{govuk}. 

Domestic abuse has substantial mental health implications, with an estimated three-quarter of survivors experiencing posttraumatic stress disorder symptoms, and are significantly more likely to report symptoms of anxiety and depression compared to the general population \cite{ferrari}. The long lasting impacts of domestic abuse are not limited to the direct target of abuse. Witnessing domestic abuse at home can have severe developmental impacts on children, including an increased risk of mental and physical health problems, difficulties in interpersonal relationships in later life, worse educational attainment, and a higher likelihood of engaging in criminal behaviours \cite{callaghan}.

As the legal definition reflects, one characteristic of domestic abuse that differentiates it from other violent crimes is its repeated nature. Estimates show that before getting effective help, on average, survivors live in the abusive relationship for 2.7 years, experiencing an estimated 50 cases of abuse \cite{SafeLives2015}. The most reliable statistics on domestic abuse in the UK is the Crime Survey for England and Wales (CSEW; \citeNP{ONS}), a victimisation survey, which includes a self-completion module on domestic abuse. According to the CSEW, of those respondents who experienced any form of domestic abuse between April 2017 and March 2018, only 17\% reported it to the police \cite{ONS}. This extremely high level of underreporting is another characteristic that is specific to domestic abuse compared to other types of crimes. 


In the most extreme cases, domestic abuse can culminate in domestic homicide. In the period between April, 2017 and March, 2018, 70 people in England and Wales were killed by their current or former partner, 90\% of these victims were women \cite{homic}, demonstrating that domestic abuse is a fundamentally gendered phenomenon. While the pervasive problem of underreporting poses a significant obstacle to deriving reliable estimates of the true extent of the problem, the economic cost of domestic abuse in England and Wales between April, 2016 and March, 2017 was estimated to be \pounds 66 billion \cite{costs}. The largest component of this cost is represented by the physical and emotional consequences of abuse, reflected in a reduced expected quality of life for survivors. In addition, lost economic output resulting from missed workdays and reduced productivity, as well as costs to the health care system also significantly contribute to the overall figure. 

This research programme consists of four research projects aiming to deliver useful, policy-relevant insights about domestic abuse. First, we will explore the demographic and socio-economic characteristics of victims of domestic abuse, the reasons behind police mis-recording of domestic abuse cases, and victim's perception of the police after reporting an incident. Second, we will investigate the factors predicting suffering serious injury and the decision to report. Third, we will examine the long-lasting effects of being a victim of domestic abuse or witnessing domestic abuse at home as a child. Finally, we explore the association between exogenous, time-varying factors, such as changes in alcohol consumption and financial resources, and the prevalence of domestic abuse. 

%Recognising the severity of this widespread societal problem led the UK government to develop a strategic plan to tackle domestic abuse with a funding of £100m for the period between 2016 and 2020. The action plan focuses on a range of areas to reduce the prevalence of domestic abuse, including increasing awareness and willingness to report through education and information campaigns, introducing legal measures to increase survivor safety (e.g., Domestic Violence Protection Order, Domestic Violence Disclosure Scheme, intervention programmes), and improve the responses of support services and health care professionals \cite{vawag}. 

%These interventions are crucial in light of the fact that after a long decline, domestic violence against women has significantly increased between the period of 2008 and 2014 \textit{ref}. 

%\newpage


\section{Data sources}

The research projects detailed below draw on various data sources. All of the projects rely on the CSEW, an annual, cross-sectional representative survey collecting information on victims of crime across England and Wales, including 30,000--40,000 households every year \cite{OfficeforNationalStatistics2019}. The survey has a self-completion module containing questions about the respondent's experiences of domestic abuse throughout their life (since the age of 16), and in the past 12 months in particular. Given the serious problem of underreporting, one of the unique characteristic of domestic abuse, the CSEW is the most reliable source of information on the prevalence of domestic abuse in England and Wales. 

Some of the projects also benefit from a crime dataset provided by the West Midlands Police (WMP), which includes all crimes and incidents recorded by the police force in the period between January, 2010 and October, 2018. The WMP is the third largest police force in England, serving a population of 2.9 million \cite{Homeoffice}. Crimes that have a domestic abuse marker indicate cases of domestic abuse that meet the criteria for notifiable offences in the UK, whereas domestic abuse incidents refer to cases that do not qualify as a crime. About 31\% of all crimes and incidents have a domestic abuse marker. For each record in this dataset, we have rich information about the circumstances of the case, including exact location and time of the incident or crime, the gender, age and ethnicity of the offender and victim, \textbf{their address}, and the severity of the injury sustained by the victim, if any. The first and last occurrence of the offence, as well as the exact time of reporting is also recorded. Each person in this dataset has a unique person identifier, allowing us to follow people over time. \textbf{\emph{\textit{Sentence about access to this dataset.}}}



These two datasets have different merits, and combining them allows us to gain a deeper understanding the characteristics and dynamics of domestic abuse, and deliver policy-relevant insights. A unique benefit of the CSEW is that it provides information on a control group (those who have not suffered domestic abuse),  allowing us to investigate the causal effect of domestic abuse. In addition, the CSEW is more likely to give a better estimate of the true prevalence of domestic abuse, and contains rich demographic, socio-economic and geographic (LSOA level) information on the respondent (survivor). However, owing to the fact that it is a victimisation survey, it does not contain any information on the perpetrators, and has limited longitudinal information (respondents are asked about whether they suffered domestic abuse in the past 12 months, but the dates of those incidents are not specified). In contrast, the WMP crime dataset has information (age, gender, ethnicity) on both the victim and the offender, and records all reported incidents for the same victim-offender pair, allowing us to conduct a quantitative analysis of the dynamics of abusive relationships, and identify the predictors of escalation. Information on the exact timing of the abuse and the time of reporting will help us identify high-risk times of the year and understand the factors influencing the decision to report. Furthermore, the crime dataset provides us with information on other criminal behaviours of the offender. In all the analysis plans outlined below, we will not link the individuals in these datasets.


Our last proposed project will also rely on a transaction per row dataset containing credit card and current account spending of 20 million customers across the UK. We have access to this data as part of a data sharing agreement with a very large UK bank. As we know which neighbourhood (LSOA) the card owners live in, this data will allow us to construct a neighbourhood-level measure of spending on alcohol, gambling, and benefit payments throughout time. We will not use individual-level data or match people in this dataset with people in the WMP dataset. 


In addition, in analyses that rely on neighbourhood-level characteristics, we will use the Incides of Multiple Deprivation (IMD; \citeNP{IMD}). The IMD ranks all 32,844 LSOAs in the UK based on their relative deprivation on various domains, including income, employment, education, health, crime, housing and living environment, providing us with a complex descriptor of the socio-economic characteristics of each neighbourhood in the country.  



\newpage

\section*{Research Programme}

This research programme involves four research projects exploring different aspects of domestic abuse. Drawing on the extensive, rich respondent information in the CSEW, and the unique, temporal nature of the WMP, we aim to complement, extend and deepen our understanding of domestic abuse. In outlining the analysis plan for each of these projects, we will specify a set of statistical analyses we plan to conduct for the sake of clarity (including dataset, unit of analysis, outcome and explanatory variables, and statistical method). We will preregister all of our analysis plans using the Open Science Framework (OSF) ahead of beginning research. We will aim to complement all analyses with robustness checks, and further exploratory regressions as the data allows (see, \textbf{\textit{Trendl Football}}, for an example of this approach).

\subsection*{1. Understanding domestic abuse: survivor characteristics and police mis-recording}

 \textcolor{red}{\textit{Short intro about aims, highlighting added value.}}
In this research project, we will first use the CSEW to investigate the demographic and socio-economic characteristics of survivors of domestic abuse. Exploring the risk factors in domestic abuse victimisation is key in designing effectively targeted policy measures. We will also use the CSEW to explore the police's response to reported cases of domestic abuse. Increasing victim's trust in the police through improving law enforcement response to reported incidents is key in encouraging victims to report domestic abuse and prevent further harm.

%Second, we will estimate the true extent of domestic abuse victimisation. Given the multifaceted nature of domestic abuse, and the pervasive problem of underreporting, it is very hard to reliably estimate the prevalence of domestic abuse. We wish to do so by combining the strengths of the two datasets (that is, the CSEW provides a more reliable measure of the true rate of victimisation across different offence types, while the WMP provides unique information on whether various types of offences were domestic abuse-related). Estimating the true prevalence of domestic abuse across various types of criminal offences will help us to obtain a better estimate of the societal cost associated with domestic abuse.

%Exploring the risk factors in domestic abuse victimisation, and the true extent of the problem of underreporting is crucial for designing effective interventions to reduce the risk of escalation that may culminate in domestic homicide. 

%Finally, we will estimate the extent of police mis-recording of domestic abuse. Increasing victim's trust in the police through improving law enforcement response to reported incidents is key in encouraging victims to report domestic abuse and prevent further harm. By contrasting estimated levels from reported incidents within the West Midlands from the two datasets, we can identify any reporting gaps between the two sources of data that might arise from police mis-recording.

%\textcolor{red}{\textit{What we know so far}} 
%
%7.9\% women and 4.2\% men reported to have experienced domestic abuse in the past year; women between 20-24 are most likely to have experienced domestic abuse; for men, age doesn't represent a significant risk factor;  divorced or separated status are significant risk factors across both sexes; long-term illness or disability is a significant predictor across both sexes; single parent households are the most vulnerable; women in the lowest household income bracket are the most likely to have experienced domestic abuse, the same is true for men but the pattern is less pronounced; only 17\% of people who have experienced domestic abuse in the past year have reported to the police, there is no significant difference in this proportion across genders; bisexual women are twice as more likely to experience DA compared to heterosexual women; mixed/multiple ethnicity; social housing; unemployment.   
%single mothers and mothers with young children are more at risk
%
%Aim 3. \href{https://www.justiceinspectorates.gov.uk/hmicfrs/news/news-feed/little-evidence-of-improvements-in-crime-recording-at-west-midlands-police/}{HMICFRS about West Midlands mis-recording}

\textcolor{red}{\textit{Detailed plan}} First, using the CSEW data, we will extend our understanding of the demographic and socio-economic factors predicting domestic abuse victimisation. The CSEW provides a broad range of information on the characteristics of the respondent, including age, sex, marital status, number of children, ethnicity, education, employment, income, benefit history, physical and mental health, frequency of going out, house and car ownership, self-reported well-being, and frequency of drug and alcohol use. Previous research has identified a number of risk factors predicting domestic abuse victimisation (being female, young, unemployed, separated or divorced, living in single parent household and earning in the lowest income bracket), based on descriptive statistics derived from the CSEW data \cite{ONS}. Building on what we know so far, we will explore the predictive power of a broader range of victim characteristics in statistical model, using a logistic regression approach. Investigating the explanatory power of these characteristics within one statistical model will provide us with a deeper understanding of the importance of these factors in predicting domestic abuse victimisation, compared to a purely descriptive statistics approach. We are also interested in how the predictive power of victim characteristics vary with the type of abuse suffered (physical abuse, threats, sexual abuse).

 However, given the complex nature of domestic abuse, it is likely that there are important interactions between various victim characteristics in predicting the risk of domestic abuse victimisation (for example, we would expect that not having stable employment is a more significant risk factor of domestic abuse victimisation for those who have young children). To uncover the structure of these interactions without having to a priori specify them, as is required in a logistic regression model, we will also use a random forest classification algorithm, a machine learning method that has the ability to detect non-linear relationships between variables, allowing us to identify particularly vulnerable subgroups of the population. Insights from both methods will provide us with the first extensive characterisation of the risk factors predicting domestic abuse victimisation in England, using a national-level, large representative sample like the CSEW.
 
 %When a crime is not recorded the force may have recorded details of the report on a different record, such as an incident log. It is these records that HMICFRS can audit to establish the extent to which reported crime is being recorded. In some cases an officer will have responded to the report.
 
 
 Her Majesty's Inspectorate of Constabulary and Fires \& Rescue Services (HMICFRS) oversees police forces in England Wales, and regularly inspects their responses to reported crime, including cases of domestic abuse. Their latest crime data integrity inspection of 32 police forces have found that 20 forces require significant improvements in their crime recording practices \cite{HerMajestysInspectorateofConstabularyandFires&RescueServices2019}. These concerns are especially pertinent in the context of domestic abuse, for example, the latest report found that the WMP have failed to record 25\% of reported crimes that were domestic abuse-related.
 
 

 We will explore whether the extent of crime mis-recording depends on victim characteristics (e.g., a crime reported by a victim with a mental illness, substance dependency or language difficulties might be less likely to be recorded as a crime by the police). Using the CSEW, we will be able to tell whether the domestic abuse reported by the respondent is likely to amount to a crime according to the list of notifiable offences \cite{countingrules}. We will run a multinomial regression with police force fixed effects for each broad crime type (sexual, physical violence, harassment and stalking), predicting whether the police's response (no action, warning, arrest, charge) depends on the socio-economic characteristics of the victim. Using a fixed effects approach will tell us the relative importance of police force and victim characteristics in crime mis-recording. To complement this analysis and explore potential interactions between our explanatory variables, we will also investigate this question with a mixed effects random forest with police force clustering. Second, using the same statistical model for each type of crime, we will also investigate how victim's satisfaction with the police, and their perception of their personal security after reporting a crime depends on the police action taken, and their socio-economic characteristics.  
 
% Using the CSEW, we are interested in understanding how for each offence type (sexual, physical with injury, stalking and harassment), the police's response depend on the victim's socio-economic characteristics 
% 
% From the CSEW, we can estimate what percentage of reported cases that ended in warning, arrest, charge, or no action, a well as the victim's satisfaction with the police handling the case and their perception of their own security as a result of the outcome.  
% 
%https://www.justiceinspectorates.gov.uk/hmicfrs/our-work/article/crime-data-integrity/reports-rolling-programme-crime-data-integrity/
   

 
 
% Police forces in England, and particularly the West Midlands Police, have previously been criticised for mis-recording violent cases of violent crime, including \href{https://www.bbc.co.uk/news/uk-england-468676578}{domestic abuse.} Mis-recording these already heavily underreported crimes decrease victim's trust in the Criminal Justice System and their willingness to report subsequent incidents. \textit{Why don't we use the aggregate number of \href{https://www.ons.gov.uk/peoplepopulationandcommunity/crimeandjustice/datasets/domesticabuseinenglandandwalesappendixtables}{police-recorded domestic abuse cases for each police force} to do this? But if we don't particularly rely on the WMP data is there a point in this? They use \href{https://www.ons.gov.uk/peoplepopulationandcommunity/crimeandjustice/bulletins/domesticabuseinenglandandwales/yearendingmarch2018}{CSEW and police data} in this report, but they don't seem to compare the two}

\begin{table}[!htbp]
\caption{Understanding domestic abuse: survivor characteristics and police mis-recording, analysis plan}
  \begin{threeparttable}[t]
  \centering
       \begin{tabular}{ L{2.8cm}  C{1.8cm}  C{1.8cm}  C{3.2cm}  C{2.8cm}  C{2cm} }
    \toprule
     \textbf{Research question} & \textbf{Dataset}    & \textbf{Unit of analysis} & \textbf{Outcome variables} & \textbf{Explanatory variables} & \textbf{Model} \\
    \midrule
    What are the characteristics of domestic abuse victims? & CSEW & Individual (respondent-level) & Domestic abuse victimisation; Victimisation by type (physical, sexual abuse, threats) & demographic and socio-economic characteristics & Logistic regression/random forest \\
        \midrule
    How does the police response depend on victim characteristics? & CSEW (reported domestic abuse, by type of offence) & Individual (respondent-level) & Police action taken (no action, warning, arrest, charge) & demographic and socio-economic characteristics of the victim & Multinomial logistic regression with  fixed effects/ mixed effects random forest \\
     \midrule
    What predicts victims' trust in the police? & CSEW (reported domestic abuse, by type of offence) & Individual (respondent-level)& Satisfaction with police action, perception of security after reporting & police action taken, demographic and socio-economic characteristics of the victim & Multinomial logistic regression with fixed effects/ mixed effects random forest \\
%    Beznau II & 23.10.1971 & Pressurised water reactor & 365   & 2031 \\
%    Mühleberg & 01.07.1971 & Boiling water reactor & 373   & 2019\\
%    Gösgen & 02.02.1979 & Pressurised water reactor & 1010  & 2039 \\
%    Leibstadt & 24.05.1984 & Boiling water reactor & 1220  & 2044 \\
     \bottomrule
  \end{tabular}
%     \begin{tablenotes}
%     \item[1] Assumption: life time of 60 years.
%     \item[2] Official shut-down.
%   \end{tablenotes}
    \end{threeparttable}%
  \label{tab:addlabel}%
\end{table}%
\FloatBarrier
%Second, we can validate some of these findings using the WMP data, since the CSEW contains information on whether the survivor reported the domestic abuse to the police. The WMP contains information about the sex and age of victims. We can contrast the sex-age victimisation numbers from the WMP with corresponding population estimates to validate the findings of the CSEW \textit{(we were initially talking about using victims of burglary, but maybe this is better?)}. We can also complement this analysis with data on the offender's sex  and age, to estimate the prevalence of different forms of reported domestic abuse (child to parent violence, parent to child violence, intimate partner violence) and contrast it with estimates based on self-reported data from the CSEW, which distinguishes between family and intimate partner violence.

%As mentioned, domestic abuse is a complex phenomenon that encompasses a number of notifiable criminal offences that are usually repeat incidents in nature (e.g., violence against the person including physical or verbal abuse, sexual offences, criminal damage, stalking and harassment). However, the CSEW mainly focuses on 




%\textbf{Aim 2.}  The CSEW is used to estimate the costs of domestic abuse in the UK. While a unique strength of the CSEW is that it provides reliable estimates of the proportion of the population who have been a survivor of a certain crime, it cannot capture the true extent of domestic abuse victimisation for several reasons. First, because it does not collect detailed information on various aspects of repeat domestic abuse victimisation (an inherent characteristic of domestic abuse cases), as it only records the number of incidents involving physical violence perpetrated by household members (and ignores repeated cases of non-physical violence). In addition, survivors who suffer frequent, ongoing abuse are less likely to accurately remember the exact number of incidents. Second, because it only focuses on certain sub-categories of domestic abuse (physical violence, sexual assault, stalking), while ignores others (criminal damage, verbal abuse, financial abuse, burglary). In addition, it does not collect information on whether different offences against the same survivor were perpetrated by the same offender. In contrast, the WMP data allows us to tell if a reported case was domestic abuse related (as the police are required to record this information), regardless of the actual offence committed. In addition, it allows us to identify various incidents with the same victim-offender pair, providing us with reliable data on reported cases of repeat victimisation. We can estimate the co-morbidity of domestic abuse victimisation with other types of criminal victimisation using the CSEW. We can combine this with the estimated likelihood of domestic abuse victimisation by offence type to get a better estimate of the overall extent of domestic abuse victimisation (reported and unreported). \textit{is this silly? then we can maybe estimate what percentage of true victimisation is domestic-abuse related? the £66 billion cost estimate is based on the CSEW}


%It is common that individuals are more likely to be a survivor of crime if they have previously been a survivor of crime, and if they have perpetrated crimes. It is also more likely that an individual will commit criminal activities if they have done so before, and if they have previously been the survivor of crimes. We will use our datasets to improve our understanding of the temporal patterns of victims and perpetrators, particularly around domestic abuse. The high level of detail in the CSEW allows us to examine the comorbidity of domestic abuse with other crimes during the previous year, and since the age of 16. 


%By contrasting extrapolated estimates from the CSEW (which is a representative sample) with WMP data on the recorded number of reported cases, we can estimate the true extent of the mis-recording of domestic abuse cases by the WMP. Given the repeated cross-sectional nature of the CSEW, and the longitudinal nature of the WMP, we can quantify the proportion of mis-recorded incidents over time. We can complement this investigation with analysing the self-reported outcome of incidents that have been reported to the police from the CSEW, and the survivor's perception of the CJS.



%co-occurrence of other types of crimes - we can contrast results from the CSEW with results from our crime data - e.g., if we see that those reporting domestic abuse have also experienced criminal damage in the CSEW, we can check in the WMP data if criminal damage or home theft is usually a spillover effect from domestic abuse ; deliberate violence; account theft



%We will use the CSEW to understand the demographic and socioeconomic predictors of the decision to report the abuse to the police, and identify specific groups who are less likely to report. We are able to do this because CSEW respondents indicated whether they reported the abuse to the police  and whether the police came to know of the abuse in another way (Question X). 
%
%Regression 1: data --- subset of people who in CSEW responding that they have been abused
%
%	abuse or not ~
%   	[demographics] age + sex + marital status + earnings + ...
%	other crime dummies + ...
%
%
%	unreported abuse / reported abuse ~ 
%	[demographics] age + sex + marital status + earnings + ... 
%	other crime dummies + ...
%	[information about abuse] level of harm + ...
	

%Respondents also answered questions indicating the level of harm that they have suffered, which will allow us to estimate how reporting and police awareness depend upon the level of harm. \NS{Anna: broaden this paragraph to be about all of the covariates}

%Linear probability model (and logit model?)

%We can contrast and validate these findings using the WMP dataset. 
%The WMP data are cases that have been reported to the police. The CSEW data allow an estimate of the true underlying prevalence of domestic abuse. The discrepancy between the two datasets allows us to estimate the reporting gap. 
%That is, we can use the WMP dataset to provide a second estimate of the degree of underreporting, indpendent of the CSEW respondent's indications of report. Discrepancy between reporting prevelence estimated from CSEW and the number of cases in the WMP data give important information about the police's recording and categorisation of domestic abuse cases---do police flag all of the cases of domestic abuse?  https://www.bbc.co.uk/news/uk-england-46867657

%While the CSEW provides us with accurate information on the characteristics of victims, it does not contain information on the perpetrators. Combining the two datasets helps us find commonalities between the perpetrators of abuse cases that are.


%Domestic abuse doesn't discriminate---well we can estimate this by comparing the 7.7\% for females and 4.4\% for males in the past year with the size of the coefficients.

%We can describe who is abused and their demographics in CSEW, CSEW who report, and in WMP.

%We can also compare those abused in WMP data with any other reference group in the WMP data. (e.g., burgled) and then use DA-burgled in CSEW vs WMP to back out stuff.


\textcolor{red}{\textit{Core outcomes}} 

The first extensive investigation of the demographic and socio-economic predictors of domestic abuse victimisation. A deeper understanding of the characteristics of police mis-recording and its effect on victim's trust in the CJS.

\newpage


\subsection*{2. Predicting serious harm and understanding the decision to report}

 \textcolor{red}{\textit{Short intro about aims, highlighting added value.}} In this project, we will explore the victim and offender characteristics that predict serious harm and the factors determining the decision to report the abuse. Identifying high-risk victim-offender pairs is key in preventing serious harm. Understanding what affects the decision to report is particularly important, as underreporting represents the biggest obstacle to effectively tackling domestic abuse. Insights from these analyses can inform decisions about the optimal timing and target audience for domestic abuse awareness campaigns.
%
%\textcolor{red}{\textit{What we know so far}}
%
%\href{https://www.thehotline.org/2018/09/28/escalation/}{Escalation} often happens when the abuser feels like they are losing control (e.g., because the survivor threatens to leave); separation is an important risk factors of serious harm; contact over children
%
%\href{http://www.safelives.org.uk/policy-evidence/about-domestic-abuse/how-long-do-people-live-domestic-abuse-and-when-do-they-get}{Reporting} 
%
%decision to report: serious violence; friends, family, neighbours notice; happens in public, in front of people  

\textcolor{red}{\textit{Detailed plan}} Drawing on the relative strengths of the CSEW and the WMP datasets, we will explore the relative importance of the risk factors predicting serious harm. Previous studies investigating this question mostly relied on police data and focused on identifying high-risk offenders \cite{thornton}. We will complement and extend this research by simultaneously considering the victim characteristics that are predictive of suffering serious injuries using the CSEW. Specifically, using a multinomial logistic approach, we will explore the extent to which financial independence and resilience (income bracket), mental health (self-reported well-being), drug and alcohol dependence (frequency of usage), social isolation (self-reported frequency of going out, living alone or with the perpetrator, family disputes), recent separation, and feeling frightened (increased home security, reporting feeling scared) are predictors of various levels of harm (threats, minor or serious injury). We will know the level of harm suffered, because the CSEW asks whether the respondent needed to see a doctor, nurse or other health worker, or needed to take time off work because of the abuse. Given the cross-sectional nature of the CSEW data, this analysis will be exploratory rather than causal.

Using the WMP data, we will also explore the risk factors associated with the perpetrator, using each perpetrator-victim pair as a unit of analysis. The strength of the WMP data is that it has a temporal dimension, and we can follow the same perpetrator-victim pair over time. Our detailed crime dataset will allow to investigate the extent to which the perpetrator's previous recorded self-harm and suicide attempts, previous violent and drug offences, breaches of court orders and separation from the victim will predict causing serious harm. From the detailed description of the incident, we will know if the perpetrator and the victim live together. We will explore this question using a multinomial logistic regression analysis with perpetrator-victim fixed effects, where we will use previously known information about the perpetrator to predict serious harm. Using a fixed-effects regression approach will allow us to disentangle the relative importance of our explanatory variables and perpetrator-victim specific factors in predicting serious harm.  

%Second, using the WMP data, we can analyse the temporal patterns of reported domestic abuse incidents by following  23\% of the 182,646 victim-offender pairs over an eight-year-long period. 


\begin{table}[!htbp]
\caption{Predicting serious harm and understanding the decision to report, analysis plan}
  \begin{threeparttable}[t]
  \centering
       \begin{tabular}{ L{2.8cm}  C{1.8cm}  C{1.8cm}  C{3.2cm}  C{2.8cm}  C{2cm} }
    \toprule
     \textbf{Research question} & \textbf{Dataset}    & \textbf{Unit of analysis} & \textbf{Outcome variables} & \textbf{Explanatory variables} & \textbf{Model} \\
    \midrule
    What victim characteristics predict serious harm? & CSEW & Individual (respondent-level) & Level of harm resulting from domestic abuse victimisation (no physical harm, minor, serious) & financial resources, mental health, dependence, social isolation, separation, feeling scared & Multinomial logistic regression \\
                \midrule
                    What perpetrator characteristics predict serious harm? & WMP & Perpetrator-victim pairs & Violent offences resulting in injury & previous self-harm and suicide attempts, previous violent and drug offences, breaching court orders, separation & Multinomial logistic regression with perpetrator-victim fixed effects \\
                \midrule
     What time-invariant (victim) characteristics predict the decision to report? & CSEW & Individual (respondent-level) & Reported abuse to the police & severity/type of abuse, trust in CJS, social ties, financial status, children, mental health, dependencies & Logistic regression \\
       \midrule
          What time-varying characteristics predict the decision to leave? & WMP & Perpetrator-victim pairs & Reported ongoing abuse to the police & time of year, specific holidays, length of abuse, severity of last occurrence & Logistic regression \\
     \bottomrule
  \end{tabular}
    \end{threeparttable}%
  \label{tab:addlabel}%
\end{table}%


%We can examine whether a recorded survivor of domestic abuse was previously recorded as a survivor in other crimes, and if they were subsequently a survivor of other criminal behaviours. Identifying risk factors is crucial for police services who want to identify individuals at risk of being future victims, so that they can target interventions and give the best advice possible to reduce an individual's risk. In addition, we can measure the relationship between experiencing domestic abuse, and the likelihood of being recorded as a suspect in other crimes. It is possible that some criminal behaviour is a precursor to experiencing domestic abuse whilst others follow it. 


We will also use both datasets to explore the predictors of the decision to report and escape the abuse. We are not aware of previous research exploring this question. Using the CSEW and logistic regression approach, we will investigate how the decision to report (or the decision to seek any form of external help) depends on the individual-level, time-invariant characteristics of the survivor, controlling for the severity and type of the abuse suffered. We are especially interested in exploring how trust in the Criminal Justice System (CJS), social ties (living with family, going out, length of living in the area, member of Neighbourhood Watch), financial independence (own income, employment status, car ownership), number of children, mental health, and alcohol and drug dependency affect the decision to report. Using the WMP data, we will explore time-varying factors affecting the decision to leave the abuser. Due to the high levels of underreporting, it is hard to distinguish between time periods when the number of domestic abuse incidents increase, and when the victims have a higher propensity to finally report the ongoing abuse. Here we are mainly interested in the latter case (and will explore the first case in more detail in the last research project), therefore we will focus on the subset of incidents where the victim for the first time reported a long history of abuse. We are interested in whether these reports are more likely to occur after certain days of the year (e.g., birthday of the victim or perpetrator, Christmas, Halloween, Easter, etc.), potentially reflecting cases where the abuse has suddenly escalated, or became known to outsiders.



\textcolor{red}{\textit{Core outcomes}}

A deeper understanding of the victim and perpetrator characteristics that predict serious harm, and the time-invariant and time-varying factors affecting the decision to report the abuse to the police. 

%Aim:
\FloatBarrier
%The CSEW respondents indicated the level of harm, and we can use the CSEW to estimate the cross-sectional association between the most serious harm and other comorbid factors like alcohol, other crimes, 

%Regression 3: type of abuse ~ 
%	visit hospital ~ 
%	mental health ~ 

%Separation is one of the biggest predictors of abuse resulting in serious injury. \textit{Subject to whether we can get data on the DVPOs (Domestic Violence Protection Order) and DVPNs (Domestic Violence Protection Notice) issued in the West Midlands in the relevant period, we could explore the offender characteristics that predict the likelihood of sustained perpetration after separation, and assess the overall efficacy of these new preventive measures. If not, we can still use info on location + breaches of non-molestation orders?}

%But cross-sectional data present a particular problem for making casual claims: For example, incidents involving drugs or alcohol intoxication: some individuals may end up in an abusive relationship because of an existing dependency, or a dependency may develop as a counterproductive strategy of coping with the abusive relationship. 



%Should you be identifying the couples most at risk of serious harm from time-invariant properties, or should you be following couples over time and using time-varying properties to spot when to intervene? 

%The WMP dataset will further allow us to follow victim-offender pairs over time, and explore how the severity and nature of reported abuse changes over time. 
%In the WMP data we can follow X victim-perp pairs for X years, with each pair appearing X times on average in the data.

%Regression 4: offense classification[proxy for harm] ~ .
%	type of abuse ~ .
%
%Future survivor behaviours. Future perpetrator behaviours
%
%serious event at t ~ stuff at t-1  + perpetrator previous crimes
% and 	a model with pair fixed effects
%
%Murders are very rare, so we are adopting a ``near miss'' airline / nuclear industry strategy of predicting the more prevalent but less serious incidents. 
%
%Say pair fixed effects control for all non-time varying characteristics. With and without pair fixed effects let's use see how much state vs trait matters for prediction. 


\subsection*{3. The long-lasting effects of domestic abuse}


 \textcolor{red}{\textit{Short intro about aims, highlighting added value.}}
Domestic abuse has long-lasting adverse effects on victims and those close to them. In this research project, our is aim to explore some of these consequences on the direct victims of domestic abuse and the children who live in the same household. Gaining a deeper understanding of the tangible, far-reaching consequences of domestic abuse will help to design survivor support programmes and quantify the societal harm caused by it.


%\textcolor{red}{\textit{What we know so far}}
%\href{https://www.womensaid.org.uk/information-support/what-is-domestic-abuse/the-nature-and-impact-of-domestic-abuse/}{Women's aid, effect on victims} increased likelihood of mental health problems (isolation, low self-esteem), alcohol and drug dependency, financial consequences due to financial abuse or losing a job as a result of the abuse, fighting back?
%
%\href{https://uksaysnomore.org/effects-of-domestic-abuse-on-children/}{Effect on children I} witnessing violence is major predictor for PTSD and depressive symptoms in children; behavioural problems, bullying (either victims or perpetrators)
%
%\href{https://www.psychologytoday.com/gb/blog/progress-notes/201902/alarming-effects-childrens-exposure-domestic-violence}{Effect on children II} cognitive development is affected, learning difficulties are more common, difficulties in social relationships, withdrawal, antisocial behaviours, they are more likely to become both perpetrators and victims of domestic abuse; protective factors: literacy, intelligence, social competence, supportive relationship with at least one influential adult 


\textcolor{red}{\textit{Detailed plan}} First, we will explore the causal effect of past experiences of domestic abuse on present socio-economic outcomes. First, using the CSEW, we will explore how witnessing abuse at home in childhood affects educational, employment and health outcomes in adulthood. The 2016 CSEW included a module asking respondents about whether they witnessed domestic abuse at home as a child. Descriptive statistics of this dataset revealed that those who reported to have witnessed abuse as a child were significantly more likely to report long-term health problems, and live in a single parent household as an adult \cite{ONSChildhood}. However, due to the possibility of various confounders (e.g., those growing up in economically deprived areas are more likely to witnessing abuse, but economic deprivation also affects educational outcomes), these descriptive statistics do not reveal the causal effect of witnessing abuse as a child on socio-economic outcomes in adulthood. To overcome this difficulty in quantifying this causal relationship, we will use propensity score matching, a statistical approach that allows for causal inferences to be drawn about the effect of witnessing abuse on educational, employment, and mental and physical health outcomes, and the probability of experiencing domestic abuse in adulthood. In a similar vein, using propensity score matching and the CSEW, we will explore how past experiences of domestic abuse affect current health outcomes, by comparing people who have experienced domestic abuse in the past, but not in the last 12 months, with people who have never experienced domestic abuse, but otherwise have similar socio-economic characteristics.


Second, using both the CSEW and WMP, we will explore how children's behavioural problems are affected by witnessing domestic abuse at home. Using data from the 10-15 year old questionnaire of the CSEW (which can be linked to the adult questionnaire), we can examine the effect of domestic abuse on children living in such households. The children's questionnaire records information about experiences with bullying, carrying knives, gang membership, school truancy, learning difficulties and health outcomes, including drug use and drinking behaviour. Using this dataset and a propensity score matching approach, we will explore the causal effect of living in an abusive household on childhood behavioural outcomes.

We will complement this analysis using data on young offenders from the WMP. We will identify if young offenders' home addresses can be linked with a domestic abuse incident which will allow us to explore whether the type of domestic abuse in the household (if any) predict the type of crime they will engage in (property-related, public order offence, violent), controlling for the socio-economic characteristics of the neighbourhood (from the Indices of Multiple Deprivation; IMD).



\begin{table}[!htbp]
\caption{The long-lasting effects of domestic abuse, analysis plan}
  \begin{threeparttable}[t]
  \centering
       \begin{tabular}{ L{2.8cm}  C{1.8cm}  C{1.8cm}  C{3.2cm}  C{2.8cm}  C{2cm} }
    \toprule
     \textbf{Research question} & \textbf{Dataset}    & \textbf{Unit of analysis} & \textbf{Outcome variables} & \textbf{Explanatory variables} & \textbf{Model} \\
    \midrule
    How does witnessing domestic abuse as a child affect socio-economic outcomes in adulthood? & CSEW & Individual (respondent-level) & Educational attainment, employment status and health-related outcomes, including domestic abuse victimisation & Witnessing domestic abuse & Propensity score matching \\
                \midrule
                   How does experiences of past (but not recent) domestic abuse affect socio-economic outcomes in adulthood? & CSEW & Individual (respondent-level) & Current health and employment outcomes & Degree and type of domestic abuse suffered & Propensity score matching \\
                \midrule
     How does living a household with domestic abuse affect behavioural outcomes in childhood? & CSEW & Individual (respondent-level) & Bullying, involvement with gangs, school truancy, carrying knives, drugs and alcohol usage, learning difficulties, health & Degree and type of domestic abuse in household & Propensity score matching \\
                \midrule
          Are young offenders from abusive households more violent? & WMP, IMD & Young offenders (individual) & Type of offence (property-related, public order offence, violent) & Household LSOA socio-economic characteristics, abusive household or not & Multinomial logistic regression \\
     \bottomrule
  \end{tabular}
    \end{threeparttable}%
  \label{tab:addlabel}%
\end{table}%

%Using data from the CSEW and well as population data of the West Midlands, we can estimate how many children and young people below 18 in each LSOA of the West Midlands are at risk of witnessing domestic abuse at home each year. 

%\textit{The problem is that we don't know the address of the young offenders. Could we the location info on where they committed the offences as a proxy of the LSOA where they live? not sure if this is viable; defining a gang member is problematic in the crime data - if something serious is committed probably only one of them will be charged with it, maybe we should just concentrate on criminal activities; I was trying to identify children who were victims of something happening at an address where DA happened and perpetrated by a DA offender - it's very narrow and excludes children who *just* witnessed abuse; however, the location id doesn't seem to specific to one place - one id can correspond to a road, dwelling, shop etc.}


%We can use school exclusion data - unfortunately this data is only available on the local authority and school-type (state-funded primary, secondary, special) level...

%There's a children's questionnaire, and they ask about gang membership. However, the sample size is very low (3,100 for the whole country). 

%of the 119,539 under 18s who were victims of crimes (domestic/child abuse, or other more serious physical/sexual violence) in the dataset, 11\% were later offenders with an average of 2 offences (mostly domestic abuse, physical violence, drug possession and theft) - these are only the kids we have evidence were traumatised at some point

%of the 149,759 da offenders, 19\% have an incident against an under 18 





%Aim:
%
%CSEW regression: child gang membership ~ any domestic abuse or reported DA [2x2 table of abuse/no abuse by gang / no gang]
%Also estimate, given gang membership what fraction are DA---which links ot WMP analysis below
%
%It also contains information about  multiple members of the household. This allows us to examine whether young individuals living in a household with a survivor of domestic abuse are more likely to be victims or perpetrators of other types of crime. By using other questions in the CSEW, we can also measure whether these children are more likely to perpetrate crimes, and to be involved in gang related activities.
% 
% explore the lasting effect of domestic abuse on the victims and their children (do we see kids having behavioural problems? No control group unfortunately)- CSEW question about gang membership - are those who report having a child gang member more likely to be victims of domestic abuse? Following people over time (Dynamic topic modelling for victims)- what’s the temporal order? 
%
%Match on name, location, and date of birth in WMP data to see effect on children
%
%WMP regression:
%
%take every juvinile in a gang and try to match all to domestic abuse cases----estimates of what fraction of gang members have co-morbid domestic abuse
%
%What fraction of reported child abuse has comorbid DA; what fraction of DA has comorbid child abuse
%
%Using school census we can go futher and estimate frequencies in a 2x2 child abuse  by domestic abuse table---providing we can match children to mothers.

\textcolor{red}{\textit{Core outcomes}}
 The first UK-based quantitative exploration of the causal effect of domestic abuse and childhood abuse on socio-economic outcomes.
\FloatBarrier
\subsection*{Environmental factors}

\textcolor{red}{\textit{Short intro about aims, highlighting added value.}}
In this research project, we will be using the CSEW and WMP datasets, as well as external sources of data to investigate some of the environmental predictors of domestic abuse. We will first explore how neighbourhood-level characteristics affect domestic abuse victimisation and the willingness to report. Understanding how neighbourhood-level characteristics affect the prevalence of domestic abuse and reporting behaviour is crucial for the most effective distribution of resources to help victims and targeted awareness campaigns. 

We will also explore how time-varying, exogenous factors affect domestic abuse. We will first explore the effect of local events that increase alcohol-consumption, since our previous study have found a significant relationship between alcohol-related domestic abuse and football. Second, we will explore the relationship between financial stress and the prevalence of domestic abuse. First, we will explore the association between gambling accessibility in the local area, and reported numbers of domestic abuse. Second, we will examine how the roll-out of the Universal Credit system affected levels of domestic abuse across the West Midlands. Apart from delivering important policy-relevant insights, these investigations will also aim to inform the local police about the optimal allocation of police resources. 



%\textcolor{red}{\textit{What we know so far}} interestingly, there are no studies about this using UK data; neighbourhood-level predictors of domestic abuse: social disorganization theory (socio-economic disadvantage and residential instability disrupt social bonds a limit collective ability to maintain control and recognise IPV); cultural norms through social learning process (high IPV neighbourhoods where people observe influential others being rewarded for engaging in IPV, whereas in low IPV neighbourhoods perpetrators get socially ostracized); women's empowerment reduces IPV, especially financial independence;  abuser's lack of employment and alcohol consumption can be a risk factor; alcohol outlet density; churches, playgrounds, community networks might foster community cohesion and reduces IPV;  however, in deprived neighbourhoods, where IPV is accepted social interactions can increase IPV
%
%\href{https://www.sciencedirect.com/science/article/abs/pii/S0277953616302891}{The relationship between electronic gaming machine accessibility and police-recorded domestic violence: A spatio-temporal analysis of 654 postcodes in Victoria, Australia, 2005–2014}
%
%\href{https://journals.sagepub.com/doi/abs/10.1177/1525107115623938}{Economic Stress and Domestic Violence: Examining the Impact of Mortgage Foreclosures on Incidents Reported to the Police }
%
%\href{https://journals.sagepub.com/doi/abs/10.1177/1524838012445641?journalCode=tvaa}{
%The Impact of Neighborhoods on Intimate Partner Violence and Victimization
%}
%
%the type sof social tie matters: friendship ties reduce the risk of IPV, family ties have no effect
%
%\href{https://journals.sagepub.com/doi/abs/10.1177/1524838013515758}{Neighborhood Environment and Intimate Partner Violence: A Systematic Review }

\textcolor{red}{\textit{Detailed plan}} Using the CSEW, we will first explore the association between the environmental characteristics of the neighbourhood, the prevalence of domestic abuse and willingness to report. Previous literature has suggested various pathways, including reduced collective efficacy of neighbourhood, and normalisation and acceptance of violence through social norms to explain the link between neighbourhood characteristics and domestic abuse, but the question has not been explored in a UK context yet. Using a logistic regression, we will explore how the interviewer's perception of the street (signs of rubbish, vandalism, and the general condition of houses, member of Neighbourhood Watch) as well as the respondent's connection and attitude towards the neighbourhood (length of time living in the local area, noisy neighbours, rubbish lying around, teenagers hanging around, vandalism, drunken and anti-social behaviour, drug trafficking, abandoned cars, speeding traffic, police presence in the area, worries about crime levels) predict domestic abuse victimisation, after controlling for the socio-economic characteristics of the respondent. In investigating this question further, we will explore how deprivation in a neighbourhood (LSOA level, measured by the Multiple Indices of Deprivation; \citeNP{IMD}) affects the prevalence of domestic abuse (estimated from the CSEW), and the willingness to report, using a spatial regression framework. Using this statistical approach will alleviate the problem of spatial dependency in our dataset, and will allow us to derive a precise estimate of the effect. These analyses will extend our understanding of the neighbourhood-level predictors of domestic abuse victimisation, and complement our individual-level approach outlined in the first project to provide a comprehensive exploration of the factors affecting domestic abuse victimisation and willingness to report. 
 


We are also interested in exploring how time-varying external factors, including alcohol consumption, and financial stress (measured by gambling expenditure and benefit dependency) affect the propensity to engage in abusive behaviours. In investigating this question, we will benefit from a dataset containing credit card and current account spending of x customers across the UK, to which we will have access as part of a data sharing agreement with a very large UK bank. This dataset will allow us to quantify spending on alcohol over time by LSOA. By using the WMP and credit card data, we will identify time-specific changes in alcohol consumption in parts of the West Midlands by identifying externals events that may affect it (e.g., local festivals, weather, bank holidays, sport tournaments), and investigate their effect on reported alcohol-related domestic abuse (from the WMP). Previously, we have found a 60\% increase in alcohol-related domestic abuse when the England national football team won, highlighting the profound effect exogenous events can have on the propensity for violence. 


 The recent reforms of the UK benefit system, in the form of the introduction of the Universal Credit (UC) had been criticised widely, due to the temporal financial strain it imposes on the least financially resilient people in society. Drawing on previous findings about the link between financial stress and domestic abuse, we will explore what effect the roll-out of the UC across the West Midlands had on the reported number of domestic abuse cases. The fact that the exact date of the roll-out varied across the seven metropolitan boroughs within the county allows for a more precise estimation of the effect of UC on the reported number of domestic abuse incidents in the West Midlands. To further explore the effect of financial shocks on the propensity for violence, we focus on gambling. Gambling is increasingly recognised as a serious health concern across the UK, and can have an adverse impact on family finances. We will use our credit card and WMP datasets to assess the link between changes in spending on gambling and domestic abuse.

%\href{https://secure.gamblingcommission.gov.uk/PublicRegister}{List of licensed gambling businesses} We can also contact the local authorities, as they should have an up-to-date list


\begin{table}[!htbp]
\caption{Environmental factors affecting domestic abuse, analysis plan}
  \begin{threeparttable}[t]
  \centering
       \begin{tabular}{ L{2.8cm}  C{1.8cm}  C{1.8cm}  C{3.2cm}  C{2.8cm}  C{2cm} }
    \toprule
     \textbf{Research question} & \textbf{Dataset}    & \textbf{Unit of analysis} & \textbf{Outcome variables} & \textbf{Explanatory variables} & \textbf{Model} \\
    \midrule
    How do environmental characteristics of the neighbourhood predict domestic abuse? & CSEW & Individual (respondent-level) & domestic abuse victimisation/willingness to report & Perception of neighbourhood, socio-economic characteristics & Logistic regression \\
         \midrule
          How does neighbourhood deprivation predict domestic abuse? & IMD, CSEW & LSOA-level & domestic abuse victimisation/willingness to report (proportion of overall population) & Various measures of deprivation & Spatial Logistic regression \\
         \midrule
    How do temporal changes in alcohol consumption and income affect domestic abuse? & WMP, Credit card data & LSOA, day/week-level & Number of reported domestic abuse cases & Spending on alcohol and gambling, changes in benefit receipts & Poisson/Negative binomial regression \\

     \bottomrule
  \end{tabular}
    \end{threeparttable}%
  \label{tab:addlabel}%
\end{table}%

\textcolor{red}{\textit{Core outcomes}} The first extensive analysis of the neighbourhood-level predictors of domestic abuse within the UK. An exploration of the exogenous, time-varying factors, including alcohol consumption and financial stress (gambling, universal credit) affecting the reported number of domestic abuse cases in the West Midlands. 



%Unit of analysis is large geographical area in a year 
%Time span 2005-
%
%DA or not ~
%
%Given DA: report or not ~
%
%Fixed effects for region and year
%
%matched in socieoconomic
%
%
%
%WMP analysis
%
%Unit of analysis is LSOA (MSOA?) per month or per quarter
%Time span - 2010-2020
%
%Number of DA cases ~ 

%The CSEW allows us to identify the demographic and socio-economic characteristics of victims who are the least likely to report, and therefore should be in the focus of support services. We can validate and contrast these findings with insights from a spatial regression analysis of the crime data, using low-level geographical area information from the census. 

%Extend regression model X to include region fixed effects?
%
%
%\textit{A sentence about what earlier investigations have found about particular risk factors, including socio-economic deprivation.} 

%deprivation is a risk factor for women
%
%Australia study on domestic abuse and proximity to fixed-odds betting terminals, and the one on offlicences.
%
%anything that affects propensity to drink/financial stress
%
%neighbourhood characteristics (deprivation), fixed odds betting terminals, alcohol outlets, events (football, festivals)
%
%\href{https://www.birminghammail.co.uk/news/midlands-news/universal-credit-rolled-out-across-14142901}{Universal credit roll-out in the West Midlands}
%
%\href{https://www.birminghammail.co.uk/black-country/universal-credit-claimants-black-country-15946831}{Universal credit roll-out in the Black Country}
%
%
%food bank dependence? data on housing e.g. evictions? homelessness.
%
%Mention we can construct some of these variables by access to "a very large UK bank" which whom we have a data sharing agreement. We cna build them from the row-per-transactino data for spending on current account and credit cards.

\newpage

\bibliographystyle{apacite}
\bibliography{domesticabuse_refs}

\end{document}
